---
title: "dates-to-avoid"
author: "Dan Spakowicz"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(tidyverse)
```

The purpose of this script is to identify when to avoid scheduling a BEWELL Study visit 1 such that future visits do not fall on holidays. 


## Define the dates to avoid and the number weeks from that date

```{r}
dates <- c("2020-11-25", "2020-11-26", "2020-11-27", "2020-12-23", "2020-12-24",
           "2020-12-25", "2020-12-31", "2021-01-01")

ddates <- lapply(dates, as.Date)

weeks <- c(4, 6, 10)
```

Subtract the weeks from the holidays and dates neighboring holidays.

```{r}
avoid <- lapply(ddates, function(x) x - weeks(weeks)) %>%
  unlist %>%
  as.Date(., origin = "1970-01-01")

avoid
```

Combine and list 

```{r}
avoid.all <- c(as.numeric(ddates), avoid) %>%
  sort 

avoid.all.d <- 
  avoid.all %>%
  unique %>%
  as.Date(., origin = "1970-01-01")
```

# DO NOT SCHEDULE VISIT 1 on:
```{r, echo=FALSE}
avoid.all.d
```

Create a similar list focused on v3

```{r}
v3wk <- 4

avoidv3 <- lapply(ddates, function(x) x - weeks(v3wk)) %>%
  unlist %>%
  as.Date(., origin = "1970-01-01")

avoid.v3 <- c(as.numeric(ddates), avoidv3) %>%
  sort 

avoid.all.v3.d <- 
  avoid.v3 %>%
  unique %>%
  as.Date(., origin = "1970-01-01")
```
# DO NOT SCHEDULE VISIT 3 on:
```{r, echo=FALSE}
avoid.all.v3.d 
```

