---
title: "Facebook advertising"
author: "Daniel Spakowicz"
date: "9/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r}
x <- read_csv("../data/raw/2020-09-11_facebook-clicks.csv") %>%
  mutate(Date = as.Date(Date, format = "%d-%b")) %>%
  mutate(Image = "People") %>%
  rename("Cost Per Result" = "Cost per result")
```

How long did the Facebook add run before the $200 cap was reached?

```{r}
max(x$Date) - min(x$Date)
```

Did we begin to saturate the market, observable by fewer clicks per day over time?

```{r}
x %>%
  ggplot(aes(x = Date, y = `Link Clicks`)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()
```
No, it looks like the number of clicks per day remained constant throughout the campaign period.

Did the cost per click change over time, suggesting that more and more "impressions" were necessary to generate this constant level of clicks?

```{r}
x %>%
  ggplot(aes(x = Date, y = `Cost Per Result`)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()

```

```{r}
summary(lm(`Cost Per Result` ~ Date, data = x))
```

No, the cost per click did not change significantly over time.

# Review of the second ad campaign

```{r}
y <- read_csv("../data/raw/2020-10-20_facebook-clicks.csv") %>%
  select(-X5) %>%
  mutate(Date = as.Date(Date, format = "%d-%b")) %>%
  rename("Cost Per Result" = "Cost Per result")

cleanCost <- function(vec) {
  gsub("^.*\\$.*(\\d\\.\\d+)", "\\1", vec) %>%
    as.numeric
}

y2 <- 
  y %>%
  mutate_at(c("Cost Per Result", "Total Daily Cost"), cleanCost)

head(y2)
```

```{r}
z <- bind_rows(x, y2)

summary(z)
```

```{r}
z %>%
  ggplot(aes(x = Date, y = `Cost Per Result`)) +
  geom_point(aes(color = Image)) +
  stat_smooth(method = "lm", aes(color = Image)) +
  theme_bw() +
  ggsave("../figures/facebook_costperresult.png",
         height = 5, width = 8)
```

