---
title: "sample-importat-to-labii"
author: "Dan Spakowicz"
date: "1/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Goals: output a table of sample information to be input into labii for each patient visit

### Input variables
    1. SubjectID
    2. date processed
    3. time processed
    4. Freezer
    5. Freezer box urine
    6. Freezer box position urine
    7. Freezer box stool 
    8. Freezer box position stool
    9. Number of stool tub samples
    10. Visit

### Output
  A table containing the rows:
    1. 10 rows of urine samples
        5 with formic acid
        5 without formic acid
    2. 2 stool vial samples
    3. some number of stool tub samples (up to 9)
    
  Each of which has the columns: 
      Name
      SampleID
      RecordID
      Location
      Position
      Formic acid? (Yes/No)
    
```{r}
if (!require("optparse")) {
  install.packages("optparse")
  library(optparse)
}
```

Define the arguments

```{r}
option_list <- list(
  make_option(c("-r", "--record_id"), type = "character", default = 75,
              help='record ids, format is id1,id2,id3', metavar = "character"),
  make_option(c("-n", "--next_urine_box_loc"), type = "character", default = "A1",
              help = "next free location in the urine freezer box",
              metavar = "character"),
  make_option(c("-b", "--urine_freezer_box"), type = "integer", default = "1",
              help = "urine freezer box number",
              metavar = "character"),
  make_option(c("-o", "--next_stool_box_loc"), type = "character", default = "A1",
              help = "next free location in the stool freezer box",
              metavar = "character"),
  make_option(c("-c", "--stool_freezer_box"), type = "integer", default = "1",
              help = "stool freezer box number",
              metavar = "character"),
  make_option(c("-x", "--number_stool-tub-tubes"), type = "integer", default = "9",
              help = "number of tubes collected from the stool tub",
              metavar = "character"),
  make_option(c("-v", "--visit"), type = "integer", default = "1",
              help = "visit number",
              metavar = "character"),
  make_option(c("-d", "--date"), type = "character", default = Sys.Date(),
              help = "date samples were processed",
              metavar = "character"),
  make_option(c("-t", "--time"), type = "character", default = format(Sys.time(), "%H:%M"),
              help = "time samples were processed",
              metavar = "character")
)
```

Set variables for testing
```{r}
# Define the record id of the person (subject ID)
# opt$record_id <- 60

```


```{r}
opt_parser <- OptionParser(option_list = option_list)
opt <- parse_args(opt_parser)

if (is.null(opt$record_id)){
  print_help(opt_parser)
  stop("At least one record id must be supplied", call.=FALSE)
}

# Convert record_id
try({
  opt$record_id <- as.character(
    unlist(
      strsplit(opt$record_id, split = ",")
    )
  )
})
```

Create a matrix of samples

```{r}
s.fields <- c("name", "patient", "record_id", "freezer_box", "box_position", "date_processed",
              "time_processed", "project")

```

First create a table for the urine samples

```{r}
ur <- matrix(nrow = length(opt$record_id) * 10, ncol = length(s.fields))

ur <- as.data.frame(ur)

names(ur) <- s.fields

ur$record_id <- rep(opt$record_id, each = 10)
ur$record_id <- paste0("HONC60-", samp$record_id)
ur$patient <- ur$record_id

ur$freezer_box <- paste0("BEWELL-Urine Freezer Box ", opt$urine_freezer_box)

# box position
#
box_map <- paste(rep(LETTERS[1:9], each = 9), rep(seq(1:9), 9), sep = "")
start_loc <- grep(opt$next_box_loc, box_map)
ur$box_position <- box_map[start_loc:(start_loc + (nrow(samp) - 1))]

# Date
samp$date_processed <- opt$date
samp$date_processed <- as.Date(samp$date_processed, format = "%F")

# Time
samp$time_processed <- opt$time

# Project
samp$project <- opt$project

# Sample ID
#
samp$sample_id <- paste(samp$record_id, format(samp$date_processed, "%y%m%d"), seq(1:10), sep = ".")

# Name (label in Labii)
samp$name <- paste(samp$project, samp$record_id, paste(samp$freezer_box,
                                                       samp$box_position, sep = "."), samp$date_processed, sep = " ")


```

# Write output

```{r}
write.table(samp,
            file = paste(as.character(samp$date_processed[1]), opt$out_samples, sep = "_"),
            sep = "\t", row.names = FALSE, quote = FALSE)
```

    
