---
title: "on-study"
author: "Amna Bibi and Dan Spakowicz"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(lubridate)

source("00-paths.R")
```

```{r}
x <- readxl::read_xlsx(file.path(paths$box, 
                                 "participants", 
                                 "Be Well Study Schedule.xlsx"),
                       skip = 1, 
                       trim_ws = TRUE,
                       sheet = 1)

head(x)
```

```{r}
xs <- 
  x %>%
  select(`Subject ID`, `Consented date`, `Visit 1`, `Redoing V1`, `Visit 2`, `Visit 3`, `Redoing V3`, `Visit 4`, `Drop Out`) %>%
    mutate(`Visit 1` = if_else(is.na(`Redoing V1`),
                             true = `Visit 1`,
                             false = `Redoing V1`),
         `Visit 3` = if_else(is.na(`Redoing V3`),
                             true = `Visit 3`,
                             false = `Redoing V3`)) %>%
  arrange(`Consented date`) %>%
  mutate(id = seq(1, nrow(x))) %>%
  select(-`Subject ID`, -starts_with("Redoing")) 

summary(xs)
```

```{r}
xsl <- 
  xs %>%
  gather(-id, key = "event", value = "date")
```

```{r}
saveRDS(xsl, file.path(paths$derived, "study-dates.RDS"))
```

```{r}
xd <- readRDS(file.path(paths$derived, "study-dates.RDS"))
```

```{r}
end.date <- as.POSIXct(max(xs$`Visit 1`, na.rm = TRUE) + lubridate::weeks(10))

xd %>%
  ggplot(aes(date, reorder(id, -id))) +
  geom_point(aes(color = event)) +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  geom_vline(xintercept = as.POSIXct(Sys.Date()), 
             color = "red", linetype = "dotted") +
  geom_vline(xintercept = end.date, 
             color = "blue", linetype = "dotted") +
  annotate(geom = "text", 
           label = paste0("Study end = ", format(end.date, "%b %d")),
           x = end.date - weeks(16), y = max(xd$id) - 3) +
  ylab("Participant") +
  xlab("Date") +
  ggsave("../figures/on-study.png", height = 4, width = 6)
```

```{r}
xw <- 
  xd %>%
  spread(key = "event", value = "date") %>%
  mutate(on.study = (`Visit 4` - as.POSIXct(Sys.Date())) > 0,
         completed = (`Visit 4` - as.POSIXct(Sys.Date())) < 0)

t.on.study <- 
  table(xw$on.study, useNA = "ifany") %>%
  data.frame

true.on.study <- t.on.study[2, 2]
true.on.study
```

There are `r true.on.study` participants currently on study, i.e. past V1 and before V4.

```{r}
gc <- 
  xs %>%
  filter(is.na(`Drop Out`)) %>%
  select(-`Consented date`,-`Drop Out`, -starts_with("Redoing")) %>%
  gather(-id, key = "event", value = "date") %>%
  mutate(date.passed = if_else(date < as.POSIXct(Sys.Date()),
                               true = TRUE,
                               false = FALSE,
                               missing = FALSE)) %>%
  filter(date.passed == FALSE)

n.consented.on.study <- 
  gc %>%
  select(id) %>%
  unique %>%
  count

n.consented.on.study
```

There are `r n.consented.on.study$n` participants currently on study, i.e. consented and before V4.

# Number of participants who have completed the study

```{r}
t.completed <- 
  xw %>%
  count(completed)

t.completed

true.completed <- t.completed[2,2]
```
There are `r as.numeric(true.completed)` participants who have completed the study.



```{r}
left.to.complete <- 42 - (true.on.study + true.completed)
left.to.complete
```

There are `r left.to.complete` participants left to fully completed the study.

```{r}
dropout.freq <- 0.8

left.to.enroll <- left.to.complete / dropout.freq
```

With a dropout frequency of `r dropout.freq`, there are `r left.to.enroll` participants left to enroll.

# Gift cards left to purchase

How many gift cards do we expect to need to buy to complete the study?

```{r}
gc <- 
  xs %>%
  filter(is.na(`Drop Out`)) %>%
  select(-`Consented date`,-`Drop Out`, -starts_with("Redoing")) %>%
  gather(-id, key = "event", value = "date") %>%
  mutate(date.passed = if_else(date < as.POSIXct(Sys.Date()),
                               true = TRUE,
                               false = FALSE,
                               missing = FALSE)) %>%
  filter(date.passed == FALSE)

```

```{r}
amazon <- 31
walmart <- 19

nrow(gc) - (amazon + walmart)
```


# Visualizing the completeness of biospecimen collections

```{r, echo=FALSE}
y <- readxl::read_xlsx(file.path(paths$box, 
                                 "participants", 
                                 "Be Well Study Schedule.xlsx"),
                       trim_ws = TRUE,
                       sheet = 2)

head(y)
```

```{r, echo=FALSE}
z <- full_join(x, y)

head(z)
```



