---
title: "on-study"
author: "Amna Bibi and Dan Spakowicz"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)

source("00-paths.R")
```

```{r}
x <- readxl::read_xlsx(file.path(paths$box, 
                                 "participants", 
                                 "Be Well Study Schedule.xlsx"),
                       skip = 1, trim_ws = TRUE)

head(x)
```

```{r}
xs <- 
  x %>%
  select(`Subject ID`, `Consented date`, `Visit 1`, `Redoing V1`, `Visit 2`, `Visit 3`, `Redoing V3`, `Visit 4`)

summary(xs)
```

```{r}
xsl <- 
  xs %>%
  gather(-`Subject ID`, key = "event", value = "date")
```

```{r}
saveRDS(xsl, file.path(paths$derived, "study-dates.RDS"))
```

```{r}
x <- readRDS(file.path(paths$derived, "study-dates.RDS"))
```

```{r}
x %>%
  ggplot(aes(date, `Subject ID`)) +
  geom_point(aes(color = event)) +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  geom_vline(xintercept = as.POSIXct(Sys.Date()), 
             color = "red", linetype = "dotted") +
  ggsave("../figures/on-study.png", height = 3, width = 5)

```

```{r}
xw <- 
  x %>%
  spread(key = "event", value = "date") %>%
  mutate(on.study = (`Visit 4` - as.POSIXct(Sys.Date())) > 0,
         completed = (`Visit 4` - as.POSIXct(Sys.Date())) < 0)

t.on.study <- 
  table(xw$on.study, useNA = "ifany") %>%
  data.frame

true.on.study <- t.on.study[2, 2]
```

There are `r true.on.study` participants currently on study.

```{r}
t.completed <- 
  xw %>%
  count(completed)

t.completed

true.completed <- t.completed[2,2]
```
There are `r as.numeric(true.completed)` participants who have completed the study.

```{r}
left.to.complete <- 42 - (true.on.study + true.completed)
```

There are `r left.to.complete` participants left to fully completed the study.

```{r}
dropout.freq <- 0.8

left.to.enroll <- left.to.complete / dropout.freq
```

With a dropout frequency of `r dropout.freq`, there are `r left.to.enroll` participants left to enroll.

# Gift cards to purchase by Dec 14 

How many gift cards do we expect to need between now and Jan 7?

```{r}
gc <- 
  x %>%
  filter(date > as.POSIXct(Sys.Date())) %>%
  mutate(before.reoder = date < as.POSIXct("2021-01-07", tz = "EST")) %>%
  count(before.reoder)

gc
```
We will need `r gc$n[grep("TRUE", gc$before.reoder)]` gift cards between `r as.Date(Sys.Date())` and Jan 7, when we are able to place an additional order.

