---
title: "diversity"
author: "Caroline Wheeler"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(reshape2)
```

read in data w/o row names
```{r}
king <- read.csv("../data/derived/RelAbun/RelAbun_king.csv", row.names = 1)
phyl <- read.csv("../data/derived/RelAbun/RelAbun_phyl.csv",row.names = 1)
class <- read.csv("../data/derived/RelAbun/RelAbun_class.csv",row.names = 1)
ord <- read.csv("../data/derived/RelAbun/RelAbun_ord.csv", row.names = 1)
fam <- read.csv("../data/derived/RelAbun/RelAbun_fam.csv", row.names = 1)
gen <- read.csv("../data/derived/RelAbun/RelAbun_gen.csv", row.names = 1)
spec <- read.csv("../data/derived/RelAbun/RelAbun_spec.csv", row.names = 1)
```

function to run wilcox test on pre/post BRB and return table of p-values
```{r}
run_wilcox <- function(level, str1, str2){
  microbes <- colnames(level)
  tableOfResults<-data.frame(var=microbes)
  tableOfResults$p_wilcox <- NA
  rownames(tableOfResults) <- microbes

  tempR <- filter(level, grepl(str1, rownames(level)))
  tempT <- filter(level, grepl(str2, rownames(level)))

  for(mic in colnames(level)){
    x <- tempR[, mic]
    y <- tempT[, mic]
    thiscommand <- paste("thiswilcox <- wilcox.test(x, y, paired = TRUE)")
    eval(parse(text = thiscommand))
    tableOfResults[mic, "p_wilcox"] <- thiswilcox$p.value
}
  tableOfResults <- tableOfResults %>% select(p_wilcox)
  return(tableOfResults)
}
```

function to get direction of change
```{r}
get_direction <- function(df, mic){
  temp <- allMics %>%
  group_by(time) %>%
  summarize(mean(eval(as.name(mic))))
  
  dif <- as.double(temp[3,2] - temp[1,2])
  if(dif > 0){
    df[mic, "direction"] <- "increase"
  }else{
    df[mic, "direction"] <- "decrease"
  }
  return(df)
}
```

BRB all levels
```{r warning=FALSE}
spec_dif <- run_wilcox(spec, "rB", "tB")
gen_dif <- run_wilcox(gen, "rB", "tB")
fam_dif <- run_wilcox(fam, "rB", "tB")
ord_dif <- run_wilcox(ord, "rB", "tB")
class_dif <- run_wilcox(class, "rB", "tB")
phyl_dif <- run_wilcox(phyl, "rB", "tB")
king_dif <- run_wilcox(king, "rB", "tB")

#combine and filter to p-value < 0.05
BRBresults <- rbind(king_dif, phyl_dif, class_dif, ord_dif, fam_dif, gen_dif, spec_dif) %>% filter(p_wilcox < 0.05)
```

Placebo all levels
```{r warning=FALSE}
spec_dif <- run_wilcox(spec, "rP", "tP")
gen_dif <- run_wilcox(gen, "rP", "tP")
fam_dif <- run_wilcox(fam, "rP", "tP")
ord_dif <- run_wilcox(ord, "rP", "tP")
class_dif <- run_wilcox(class, "rP", "tP")
phyl_dif <- run_wilcox(phyl, "rP", "tP")
king_dif <- run_wilcox(king, "rP", "tP")

#combine and filter to p-value < 0.05
resultsPlacebo <- rbind(king_dif, phyl_dif, class_dif, ord_dif, fam_dif, gen_dif, spec_dif) %>% filter(p_wilcox < 0.05)
```

add direction to BRB results table
```{r}
BRBresults$direction <- NA

allMics <- cbind(king, phyl, class, ord, fam, gen, spec)
  
allMics <- allMics %>%
  mutate(time = rownames(allMics)) %>%
  mutate(time = substr(time, nchar(time)-1, nchar(time)))

for (microbe in rownames(BRBresults)) {
  BRBresults <- get_direction(BRBresults, microbe)
}
```

save BRBresults
```{r}
write.csv(BRBresults, "../data/derived/03-28-22_BRBmicrobiomeResults.csv")
```

add direction to Placebo results table
```{r}
for (microbe in rownames(resultsPlacebo)) {
  resultsPlacebo <- get_direction(resultsPlacebo, microbe)
}
```

save Placresults
```{r}
write.csv(resultsPlacebo, "../data/derived/03-28-22_PlaceboMicrobiomeResults.csv")
```

<!-- Top 5 -->
<!-- ```{r} -->
<!-- sigMics <- rownames(results %>% arrange(p_wilcox) %>% slice(1:6)) -->
<!-- ``` -->

boxplots of significant microbes
```{r}
sigMics <- allMics[, which((colnames(allMics) %in% rownames(BRBresults))==TRUE)]
sigMics <- sigMics %>%
  mutate(time = rownames(sigMics)) %>%
  mutate(time = substr(time, nchar(time)-1, nchar(time))) %>%
  filter(time == "rB" | time == "tB")
  
sigMics <- melt(sigMics)

bp <- 
  sigMics %>%
  ggplot(aes(x=time, y=value)) + 
  geom_violin(aes(color = time), show.legend = FALSE) +
  geom_boxplot(aes(fill=time), width = 0.1) + 
  geom_jitter(alpha = 0.2, width = 0.3) +
  xlab("") + ylab("Relative Abundance") + 
  labs(fill = "") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) + 
  scale_fill_manual(values=c("red3", "cornflowerblue")) + 
  facet_wrap(. ~ variable, scales = "free", ncol = 4)
  
bp

ggsave(filename = "../figures/aacr22_microbiomeBP.png", plot = bp, width = 12, height = 5.5, dpi = 300)
```

Summarize all microbes for visualization
```{r}
aml <- allMics %>%
  rownames_to_column("id") %>%
  separate("id", into = c("ptid", "study_visit")) %>%
  gather(-ptid, -study_visit, key = "microbe", value = "relAbun") %>%
  mutate(study_visit = fct_relevel(study_visit, c("rP", "tP", "rB", "tB"))) %>%
  mutate(relAbun = as.numeric(relAbun))


```

```{r}
aml %>%
  filter(microbe == "s__Lachnospira_pectinoschiza") %>%
  ggplot(aes(x = study_visit, y = relAbun)) +
  geom_bar(stat = "identity", aes(fill = study_visit), show.legend = FALSE) +
  scale_fill_viridis_d() +
  theme_bw() +
  labs(x = "",
       y = "Relative Abundance",
       subtitle = "Lachnospira pectinoschiza") +
  facet_wrap(~ptid)

ggsave("../figures/Lachnospira_pectinoschiza_abundance_bar_all.png",
       height = 7.5, width = 10)
```

```{r}
aml %>%
  filter(microbe == "s__Blautia_obeum") %>%
  ggplot(aes(x = study_visit, y = relAbun)) +
  geom_bar(stat = "identity", aes(fill = study_visit), show.legend = FALSE) +
  scale_fill_viridis_d() +
  theme_bw() +
  labs(x = "",
       y = "Relative Abundance",
       subtitle = "Blautia obeum") +
  facet_wrap(~ptid)

ggsave("../figures/Blautia_obeum_abundance_bar_all.png",
       height = 7.5, width = 10)
```

s__Roseburia_sp_CAG_309

```{r}
plotMicrobe <- function(df, name){
  p <- 
    df %>%
    filter(microbe == name) %>%
    ggplot(aes(x = study_visit, y = relAbun)) +
    geom_bar(stat = "identity", aes(fill = study_visit), show.legend = FALSE) +
    scale_fill_viridis_d() +
    theme_bw() +
    labs(x = "",
         y = "Relative Abundance",
         subtitle = name) +
    facet_wrap(~ptid)
  p

ggsave(paste0("../figures/", name, "_abundance_bar_all.png"),
       height = 7.5, width = 10)
  
}

plotMicrobe(aml, "s__Roseburia_sp_CAG_309")
```

