---
title: "Urine metabolomics EDA"
author: "Lane Nevers"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
```

read in files 
```{r}
met <- readxl::read_xlsx("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx")
met2 <- readxl::read_xlsx("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx")

metadata4 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata5 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata6 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata7 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata8 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata9 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata10 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata11 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata12 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata13 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata14 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata15 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata16 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata17 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata18 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
metadata19 <- read_excel("../data/raw/urine-metabolomics/Dan Urine Anthocyanin 11-10-2022.xlsx", sheet = "concentration ng mL")
timepoint.key <- readRDS("../../data/derived/clinical_manuscript.rds")$samples_long
```

testing how to best manipulate data
```{r}
met2 <- met2[-c(2:23), ]
met2.transf <- met2 %>%
  gather(key = "column_ID", value = "value", 2:81)
met2.sep <- met2.transf %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-" , remove = FALSE)
```


1.1 transform data for cy glc sulf
```{r}
metadata4 <- metadata4[-c(2:37), ]
metadata4.trans <- metadata4 %>% 
  gather(key = "column_ID", value = "value", 2:81)
metadata4.sep <- metadata4.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-" , remove = FALSE)
metadata4.sep <- metadata4.sep %>%
  rename(ng.mL=value)
metadata4.sep$text <- toupper(metadata4.sep$text)
timepoint.key[6, "timepoints"] <- "HONC60-03-v3"
timepoint.key[9, "timepoints"] <- "HONC60-05-v3"
timepoint.key[87, "timepoints"] <- "HONC60-76-v1"
timepoint.key[120, "timepoints"] <- "HONC60-97-v4"
timepoint.key <- timepoint.key %>%
   separate(timepoints, into = c("Prefix", "Suffix"), sep = "(?<=HONC60)-")
timepoint.key <- timepoint.key %>%
  separate(Suffix, into = c("numeric", "text"), sep = "-")
timepoint.key$text <- toupper(timepoint.key$text)
timepoint.key$text <- gsub("\\.1", "", timepoint.key$text)
merged.meta.tpk <- merge(metadata4.sep, timepoint.key, by = c("numeric", "text"))
merged.meta.tpk <- merged.meta.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))

```

1.2 plot for cy glc sulf
```{r}
metadata4.sep$ng.mL <- as.numeric(metadata4.sep$ng.mL)
ggplot(metadata4.sep, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL", limits = c(0,340))
```
1.3 boxplot for cy glc sulf
```{r}
merged.meta.tpk$ng.mL <- as.numeric(merged.meta.tpk$ng.mL)

non_numeric_values <- merged.meta.tpk[!is.na(as.numeric(merged.meta.tpk$ng.mL)) & !is.finite(merged.meta.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```


1.4 plot for cy glc sulf highlighted (example)
```{r}
metadata4.sep.high <- metadata4.sep
metadata4.sep.high <- subset(metadata4.sep.high, numeric %in% c(21, 61, 60, 71, 77, 81, 95, 96, 62, 85))
metadata4.sep.high$ng.mL <- as.numeric(metadata4.sep.high$ng.mL)
ggplot(metadata4.sep.high, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL", limits = c(0,310))
```

2.1 transform data for cy gluA
```{r}
metadata5 <- metadata5[-c(1, 3:37), ]
metadata5.trans <- metadata5 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata5.sepe <- metadata5.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata5.sepe <- metadata5.sepe %>%
  rename(ng.mL=value)
metadata5.sepe$text <- toupper(metadata5.sepe$text)
merged.meta5.tpk <- merge(metadata5.sepe, timepoint.key, by = c("numeric", "text"))
merged.meta5.tpk <- merged.meta5.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

2.2 extract all rows that exclude 'NA' from cy gluA
```{r}
metadata5.sepe.dupe <- metadata5.sepe[complete.cases(metadata5.sepe),]
values_to_delete <- c(95, 96, 68, 61)
metadata5.sepe.dupe <- metadata5.sepe.dupe[!(metadata5.sepe.dupe$numeric %in% values_to_delete), ]
```

2.3 plot for cy gluA
```{r}
metadata5.sepe.dupe$ng.mL <- as.numeric(metadata5.sepe.dupe$ng.mL)
ggplot(metadata5.sepe.dupe, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
2.4 graph for cy glcA
```{r}
merged.meta5.tpk$ng.mL <- as.numeric(merged.meta5.tpk$ng.mL)

non_numeric_values <- merged.meta5.tpk[!is.na(as.numeric(merged.meta.tpk$ng.mL)) & !is.finite(merged.meta5.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta5.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta5.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```
3.1 transform data for cy hex 
```{r}
metadata6 <- metadata6[-c(1,2,4:37), ]
metadata6.trans <- metadata6 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata6.sep <- metadata6.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata6.sep <- metadata6.sep %>%
  rename(ng.mL=value)
metadata6.sep$text <- toupper(metadata6.sep$text)
merged.meta6.tpk <- merge(metadata6.sep, timepoint.key, by = c("numeric", "text"))
merged.meta6.tpk <- merged.meta6.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

```{r}
merged.meta6.tpk$ng.mL <- as.numeric(merged.meta6.tpk$ng.mL)

non_numeric_values <- merged.meta6.tpk[!is.na(as.numeric(merged.meta6.tpk$ng.mL)) & !is.finite(merged.meta6.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta6.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))

ggplot(merged.meta6.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

3.2 extract all rows that exclude 'NA' from cy hex
```{r}
metadata6.sep.dup <- metadata6.sep.dup[metadata6.sep.dup$ng.mL != "NA", ]
```

4.1 transform data for cy rut
```{r}
metadata7 <- metadata7[-c(1:3,5:37), ]
metadata7.trans <- metadata7 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata7.sep <- metadata7.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata7.sep <- metadata7.sep %>%
  rename(ng.mL=value)
metadata7.sep$text <- toupper(metadata7.sep$text)
merged.meta7.tpk <- merge(metadata7.sep, timepoint.key, by = c("numeric", "text"))
merged.meta7.tpk <- merged.meta7.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

4.2 extract all rows that exclude 'NA' from cy rut
```{r}
metadata7.sep.dup <- metadata7.sep[metadata7.sep$ng.mL != "NA", ]
```
4.3 plot for cy rut
```{r}
merged.meta7.tpk$ng.mL <- as.numeric(merged.meta7.tpk$ng.mL)

non_numeric_values <- merged.meta7.tpk[!is.na(as.numeric(merged.meta7.tpk$ng.mL)) & !is.finite(merged.meta7.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta7.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta7.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```
5.1 transform data for cy sam
```{r}
metadata8 <- metadata8[-c(1:4,6:37), ]
metadata8.trans <- metadata8 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata8.sep <- metadata8.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata8.sep <- metadata8.sep %>%
  rename(ng.mL=value)
metadata8.sep$text <- toupper(metadata8.sep$text)
merged.meta8.tpk <- merge(metadata8.sep, timepoint.key, by = c("numeric", "text"))
merged.meta8.tpk <- merged.meta8.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

5.2 extract all rows that exclude 'NA' from cy sam
```{r}
metadata8.sep.dup <- metadata8.sep[metadata8.sep$ng.mL != "NA", ]
```
5.3 graph for cy sam
```{r}
merged.meta8.tpk$ng.mL <- as.numeric(merged.meta8.tpk$ng.mL)

non_numeric_values <- merged.meta8.tpk[!is.na(as.numeric(merged.meta8.tpk$ng.mL)) & !is.finite(merged.meta8.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta8.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta8.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

6.1 transform data for cy sulf
```{r}
metadata9 <- metadata9[-c(1:5,7:37), ]
metadata9.trans <- metadata9 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata9.sep <- metadata9.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata9.sep <- metadata9.sep %>%
  rename(ng.mL=value)
metadata9.sep$text <- toupper(metadata9.sep$text)
merged.meta9.tpk <- merge(metadata9.sep, timepoint.key, by = c("numeric", "text"))
merged.meta9.tpk <- merged.meta9.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

6.2 extract all rows that exclude 'NA' from cy sulf 
```{r}
metadata9.sep.dup <- metadata9.sep[metadata9.sep$ng.mL != "NA", ]
```
6.3 plot for cy sulf
```{r}
merged.meta9.tpk$ng.mL <- as.numeric(merged.meta9.tpk$ng.mL)

non_numeric_values <- merged.meta9.tpk[!is.na(as.numeric(merged.meta9.tpk$ng.mL)) & !is.finite(merged.meta9.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta9.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta9.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

7.1 transform data for cy xyrut
```{r}
metadata10 <- metadata10[-c(1:6,8:37), ]
metadata10.trans <- metadata10 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata10.sep <- metadata10.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata10.sep <- metadata10.sep %>%
  rename(ng.mL=value)
metadata10.sep$text <- toupper(metadata10.sep$text)
merged.meta10.tpk <- merge(metadata10.sep, timepoint.key, by = c("numeric", "text"))
merged.meta10.tpk <- merged.meta10.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

7.2 extract all rows that exclude 'NA' from cy xyrut
```{r}
metadata10.sep.dup <- metadata10.sep[metadata10.sep$ng.mL != "NA", ]
```

7.3 extract all rows that include pnt 77 from cy xyrut
```{r}
metadata10.sep.dup.77 <- metadata10.sep.dup[metadata10.sep.dup$numeric == 77, ]
```

7.4 plot for cy xyrut
```{r}
metadata10.sep.dup.77$ng.mL <- as.numeric(metadata10.sep.dup.77$ng.mL)
ggplot(metadata10.sep.dup.77, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
7.5 graph for cy xyrut
```{r}
merged.meta10.tpk$ng.mL <- as.numeric(merged.meta10.tpk$ng.mL)

non_numeric_values <- merged.meta10.tpk[!is.na(as.numeric(merged.meta10.tpk$ng.mL)) & !is.finite(merged.meta10.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta10.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta10.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

8.1 transform data for ferulic acid
```{r}
metadata11 <- metadata11[-c(1:7,9:37), ]
metadata11.trans <- metadata11 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata11.sep <- metadata11.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata11.sep <- metadata11.sep %>%
  rename(ng.mL=value)
metadata11.sep$text <- toupper(metadata11.sep$text)
merged.meta11.tpk <- merge(metadata11.sep, timepoint.key, by = c("numeric", "text"))
merged.meta11.tpk <- merged.meta11.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

8.2 extract all rows that exclude 'NA' from ferulic acid
```{r}
metadata11.sep.dup <- metadata11.sep[metadata11.sep$ng.mL != "NA", ]
```

8.3 plot for ferulic acid
```{r}
metadata11.sep.dup$ng.mL <- as.numeric(metadata11.sep.dup$ng.mL)
ggplot(metadata11.sep.dup, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
8.4 graph for ferulic acid
```{r}
merged.meta11.tpk$ng.mL <- as.numeric(merged.meta11.tpk$ng.mL)

non_numeric_values <- merged.meta11.tpk[!is.na(as.numeric(merged.meta11.tpk$ng.mL)) & !is.finite(merged.meta11.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta11.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta11.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

9.1 transform data for pca
```{r}
metadata12 <- metadata12[-c(1:8,10:37), ]
metadata12.trans <- metadata12 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata12.sep <- metadata12.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata12.sep <- metadata12.sep %>%
  rename(ng.mL=value)
metadata12.sep$text <- toupper(metadata12.sep$text)
merged.meta12.tpk <- merge(metadata12.sep, timepoint.key, by = c("numeric", "text"))
merged.meta12.tpk <- merged.meta12.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

9.2 extract all rows that exclude 'NA' from pca
```{r}
metadata12.sep.dup <- metadata12.sep[metadata12.sep$ng.mL != "NA", ]
```

9.3 plot for pca
```{r}
metadata12.sep.dup$ng.mL <- as.numeric(metadata12.sep.dup$ng.mL)
ggplot(metadata12.sep.dup, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
9.4 graph for pca
```{r}
merged.meta12.tpk$ng.mL <- as.numeric(merged.meta12.tpk$ng.mL)

non_numeric_values <- merged.meta12.tpk[!is.na(as.numeric(merged.meta12.tpk$ng.mL)) & !is.finite(merged.meta12.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta12.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta12.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

10.1 transform data for peon gluA
```{r}
metadata13 <- metadata13[-c(1:9,11:37), ]
metadata13.trans <- metadata13 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata13.sep <- metadata13.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata13.sep <- metadata13.sep %>%
  rename(ng.mL=value)
metadata13.sep$text <- toupper(metadata13.sep$text)
merged.meta13.tpk <- merge(metadata13.sep, timepoint.key, by = c("numeric", "text"))
merged.meta13.tpk <- merged.meta13.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

10.2 extract all rows that exclude 'NA' from peon gluA
```{r}
metadata13.sep.dup <- metadata13.sep[metadata13.sep$ng.mL != "NA", ]
```

10.3 extract all rows that include pnt 84 from peon gluA
```{r}
metadata13.sep.dup.84 <- metadata13.sep.dup[metadata13.sep.dup$numeric == 84, ]
```

10.4 plot for peon gluA
```{r}
metadata13.sep.dup.84$ng.mL <- as.numeric(metadata13.sep.dup.84$ng.mL)
ggplot(metadata13.sep.dup.84, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
10.5 graph for peon gluA
```{r}
merged.meta13.tpk$ng.mL <- as.numeric(merged.meta13.tpk$ng.mL)

non_numeric_values <- merged.meta13.tpk[!is.na(as.numeric(merged.meta13.tpk$ng.mL)) & !is.finite(merged.meta13.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta13.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta13.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

11.1 transform data for peon hex
```{r}
metadata14 <- metadata14[-c(1:10,12:37), ]
metadata14.trans <- metadata14 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata14.sep <- metadata14.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata14.sep <- metadata14.sep %>%
  rename(ng.mL=value)
metadata14.sep$text <- toupper(metadata14.sep$text)
merged.meta14.tpk <- merge(metadata14.sep, timepoint.key, by = c("numeric", "text"))
merged.meta14.tpk <- merged.meta14.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

11.2 extract all rows that exclude 'NA' from peon hex
```{r}
metadata14.sep.dup <- metadata14.sep[metadata14.sep$ng.mL != "NA", ]
```
11.3 graph for peon hex
```{r}
merged.meta14.tpk$ng.mL <- as.numeric(merged.meta14.tpk$ng.mL)

non_numeric_values <- merged.meta14.tpk[!is.na(as.numeric(merged.meta14.tpk$ng.mL)) & !is.finite(merged.meta14.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta14.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta14.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

12.1 transform data for peon rut
```{r}
metadata15 <- metadata15[-c(1:11,13:37), ]
metadata15.trans <- metadata15 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata15.sep <- metadata15.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata15.sep <- metadata15.sep %>%
  rename(ng.mL=value)
metadata15.sep$text <- toupper(metadata15.sep$text)
merged.meta15.tpk <- merge(metadata15.sep, timepoint.key, by = c("numeric", "text"))
merged.meta15.tpk <- merged.meta15.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

12.2 extract all rows that exclude 'NA' from peon rut
```{r}
metadata15.sep.dup <- metadata15.sep[metadata15.sep$ng.mL != "NA", ]
```

12.3 extract all rows that include pnt 66 from peon rut
```{r}
metadata15.sep.dup.66 <- metadata15.sep.dup[metadata15.sep.dup$numeric == 66, ]
```

12.4 plot for peon rut
```{r}
metadata15.sep.dup.66$ng.mL <- as.numeric(metadata15.sep.dup.66$ng.mL)
ggplot(metadata15.sep.dup.66, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
12.5 graph for peon rut
```{r}
merged.meta15.tpk$ng.mL <- as.numeric(merged.meta15.tpk$ng.mL)

non_numeric_values <- merged.meta15.tpk[!is.na(as.numeric(merged.meta15.tpk$ng.mL)) & !is.finite(merged.meta15.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta15.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta15.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```


13.1 transform data for peon sam
```{r}
metadata16 <- metadata16[-c(1:12,14:37), ]
metadata16.trans <- metadata16 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata16.sep <- metadata16.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata16.sep <- metadata16.sep %>%
  rename(ng.mL=value)
metadata16.sep$text <- toupper(metadata16.sep$text)
merged.meta16.tpk <- merge(metadata16.sep, timepoint.key, by = c("numeric", "text"))
merged.meta16.tpk <- merged.meta16.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

13.2 extract all rows that exclude 'NA' from peon sam
```{r}
metadata16.sep.dup <- metadata16.sep[metadata16.sep$ng.mL != "NA", ]
```

13.3 extract all rows that include pnt 77 from peon sam
```{r}
metadata16.sep.dup.77 <- metadata16.sep.dup[metadata16.sep.dup$numeric == 77, ]
```

13.4 plot for peon sam
```{r}
metadata16.sep.dup.77$ng.mL <- as.numeric(metadata16.sep.dup.77$ng.mL)
ggplot(metadata16.sep.dup.77, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
13.5 graph for peon sam
```{r}
merged.meta16.tpk$ng.mL <- as.numeric(merged.meta16.tpk$ng.mL)

non_numeric_values <- merged.meta16.tpk[!is.na(as.numeric(merged.meta16.tpk$ng.mL)) & !is.finite(merged.meta16.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta16.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta16.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

14.1 transform data for peon xyrut
```{r}
metadata17 <- metadata17[-c(1:13,15:37), ]
metadata17.trans <- metadata17 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata17.sep <- metadata17.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata17.sep <- metadata17.sep %>%
  rename(ng.mL=value)
metadata17.sep$text <- toupper(metadata17.sep$text)
merged.meta17.tpk <- merge(metadata17.sep, timepoint.key, by = c("numeric", "text"))
merged.meta17.tpk <- merged.meta17.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

14.2 extract all rows that exclude 'NA' from peon xyrut
```{r}
metadata17.sep.dup <- metadata17.sep[metadata17.sep$ng.mL != "NA", ]
```

14.3 extract all rows that include pnt 66 from peon xyrut
```{r}
metadata17.sep.dup.66 <- metadata17.sep.dup[metadata17.sep.dup$numeric == 66, ]
```

14.4 plot for peon xyrut
```{r}
metadata17.sep.dup.66$ng.mL <- as.numeric(metadata17.sep.dup.66$ng.mL)
ggplot(metadata17.sep.dup.66, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
14.5 graph for peon xyrut
```{r}
merged.meta17.tpk$ng.mL <- as.numeric(merged.meta17.tpk$ng.mL)

non_numeric_values <- merged.meta17.tpk[!is.na(as.numeric(merged.meta17.tpk$ng.mL)) & !is.finite(merged.meta17.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta17.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta17.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

15.1 transform data for pga
```{r}
metadata18 <- metadata18[-c(1:14,16:37), ]
metadata18.trans <- metadata18 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata18.sep <- metadata18.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata18.sep <- metadata18.sep %>%
  rename(ng.mL=value)
metadata18.sep$text <- toupper(metadata18.sep$text)
merged.meta18.tpk <- merge(metadata18.sep, timepoint.key, by = c("numeric", "text"))
merged.meta18.tpk <- merged.meta18.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

15.2 plot for pga 
```{r}
metadata18.sep$ng.mL <- as.numeric(metadata18.sep$ng.mL)
ggplot(metadata18.sep, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
15.3 graph for pga
```{r}
merged.meta18.tpk$ng.mL <- as.numeric(merged.meta18.tpk$ng.mL)

non_numeric_values <- merged.meta18.tpk[!is.na(as.numeric(merged.meta18.tpk$ng.mL)) & !is.finite(merged.meta18.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta18.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta18.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```

16.1 transform data for vanillic acid
```{r}
metadata19 <- metadata19[-c(1:15,17:37), ]
metadata19.trans <- metadata19 %>%
  gather(key = "column_ID", value = "value", 2:81)
metadata19.sep <- metadata19.trans %>%
  separate(column_ID, into = c("numeric", "text"), sep = "-", remove = FALSE)
metadata19.sep <- metadata19.sep %>%
  rename(ng.mL=value)
metadata19.sep$text <- toupper(metadata19.sep$text)
merged.meta19.tpk <- merge(metadata19.sep, timepoint.key, by = c("numeric", "text"))
merged.meta19.tpk <- merged.meta19.tpk %>%
  mutate(timepoint_meaning = case_when(
    Timepoint == "rP" ~ "Pre-placebo" ,
    Timepoint == "tP" ~ "Post-placebo" , 
    Timepoint == "rB" ~ "Pre-BRB" , 
    Timepoint == "tB" ~ "Post-BRB" , 
    TRUE ~ "unknown"
  ))
```

16.2 plot for vanillic acid
```{r}
metadata19.sep$ng.mL <- as.numeric(metadata19.sep$ng.mL)
ggplot(metadata19.sep, aes(x = text, y = ng.mL, color = factor(numeric))) + 
  geom_point() + 
  scale_y_continuous(name = "ng.mL")
```
16.3 graph for vanillic acid 
```{r}
merged.meta19.tpk$ng.mL <- as.numeric(merged.meta19.tpk$ng.mL)

non_numeric_values <- merged.meta19.tpk[!is.na(as.numeric(merged.meta19.tpk$ng.mL)) & !is.finite(merged.meta19.tpk$ng.mL), "ng.mL"]
if (length(non_numeric_values) > 0) {
  print("Warning: 'ng.mL' column contains non-numeric values.")
}
avg_data <- merged.meta19.tpk %>%
  group_by(timepoint_meaning) %>%
  summarize(avg_value = mean(ng.mL, na.rm = TRUE))
View(avg_data)
ggplot(merged.meta19.tpk, aes(x = factor(timepoint_meaning, levels = c("Pre-placebo", "Post-placebo", "Pre-BRB", "Post-BRB")) , y = ng.mL)) +
  geom_boxplot() +
  labs(x = "Timepoint Meaning")
```
strat cy glc sulf
```{r}
merged.meta.tpk.difference <- merged.meta.tpk %>%
  pivot_wider(names_from = timepoint_meaning, values_from = ng.mL)
merged.meta.tpk.difference <- merged.meta.tpk.difference %>%
  select(-`Pre-placebo`, -`Post-placebo`)
merged.meta.tpk.difference.filtered <- subset(merged.meta.tpk.difference, !grepl("rP|tP", Timepoint))
merged.meta.tpk.difference.filtered$`Post-BRB` <- as.numeric(merged.meta.tpk.difference.filtered$`Post-BRB`)
merged.meta.tpk.difference.filtered$`Pre-BRB` <- as.numeric(merged.meta.tpk.difference.filtered$`Pre-BRB`)
merged.meta.tpk.difference.filtered <- merged.meta.tpk.difference.filtered %>%
  mutate(Post_BRB_aligned = lead(`Post-BRB`), Pre_BRB_aligned = lag(`Pre-BRB`))
merged.meta.tpk.difference.filtered <- merged.meta.tpk.difference.filtered %>%
  select(-Difference)
merged.meta.tpk.difference.filtered <- merged.meta.tpk.difference.filtered %>%
  mutate(Difference = Post_BRB_aligned - `Pre-BRB`)
```

strat cy hex 
```{r}
merged.meta6.tpk.difference <- merged.meta6.tpk %>%
  pivot_wider(names_from = timepoint_meaning, values_from = ng.mL)
merged.meta6.tpk.difference[is.na(merged.meta6.tpk.difference)] <- 0

```

