---
title: "paa_microSig"
author: "Caroline Wheeler"
date: '2022-12-22'
output: html_document
---

---
title: "paa"
author: "Caroline Wheeler"
date: "2022-12-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(randomForest)
library(ggplot2)
library(ROCR)
library(gdata)
```

read in test data
```{r}
# bewell
b_taxa <- read.csv("../data/derived/bewell_taxonomy_relAbun.csv")

# fitness 
f_taxa <- read.table("/Volumes/spakowicz/fitness/data/derived/merged_metaphlan_output.txt",
                     header = T, stringsAsFactors = F)
```

samples wide to long 
```{r}
f_taxa <- gather(f_taxa, sample, relAbun, FF05466751:FF05464119, factor_key=TRUE)
```

serparate taxonomy into taxonomic levels 
```{r}
f_taxa <- f_taxa %>%
  separate(clade_name, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), '\\|') %>%
  rename(exo.ra = relAbun) %>%
  mutate(microbe = ifelse(!is.na(Species), Species, 
                          ifelse(!is.na(Genus), Genus,
                                 ifelse(!is.na(Family), Family,
                                        ifelse(!is.na(Order), Order,
                                               ifelse(!is.na(Class), Class,
                                                      ifelse(!is.na(Phylum), Phylum, Kingdom))))))) %>%
  select(sample, microbe, exo.ra)

b_taxa <- b_taxa %>%
  separate(Taxonomy, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), '\\|') %>%
  rename(exo.ra = RelAbun) %>%
  mutate(microbe = ifelse(!is.na(Species), Species, 
                          ifelse(!is.na(Genus), Genus,
                                 ifelse(!is.na(Family), Family,
                                        ifelse(!is.na(Order), Order,
                                               ifelse(!is.na(Class), Class,
                                                      ifelse(!is.na(Phylum), Phylum, Kingdom))))))) %>%
  select(sample, microbe, exo.ra)

```

define
```{r}
paa_up <- c("s__Clostridium_sp_UNK-MGS-6", "s__Firmicutes_bacterium_CAG_41", "s__Coprobacillus_cateniformis", "s__Clostridia_bacterium_UC5-1-2G4","s__Streptococcus_dysgalactiae", "s__Acidaminococcus_sp_HPA0509", "s__Roseburia_sp_CAG_303", "s__Butyrivibrio_sp_AE2032", "s__Coprobacillus_sp_29-1", "s__Coprobacillus_sp_D6")

paa_down <- c("s__Alistipes_sp_CAG_435", "s__Fusobacterium_hwasookii", "s__Leuconostoc_sp_DORA_2", "s__Paenibacillus_lentimorbus", "s__Fusobacterium_sp_oral_taxon_370", "s__Alcanivorax_hongdengensis", "s__Arthrobacter_sp_LS16", "s__Bacillus_sp_5B6", "s__Bacillus_velezensis", "s__Fusobacterium_sp_OBRC1")
```

Calculate signature score
```{r}
# Function to calculate a microbe signature given a set of up and down regulated microbes and the relAbun of sample data
#
# @sig_up character vector listing microbes in given signature that are up-regulated
# @sig_down character vector listing microbes in given signature that are down-regulated
# @taxa df containing columns 'sample', 'microbe', 'exo.ra'
#
getMicroSigScore <- function(sig_up, sig_down, taxa){
  # z score transform relAbun 
  df <- taxa %>%
    group_by(sample) %>%
    mutate(z_score = scale(exo.ra))
  
  upMics <- df %>%
    filter(microbe %in% sig_up)
  
  upTotal <- length(unique(upMics$microbe))
  
  upMics <- upMics %>%  
    group_by(sample) %>%
    summarize(tempUp = sum(exo.ra))
  
  downMics <- df %>%
    filter(microbe %in% sig_down)
  
  downTotal <- length(unique(downMics$microbe))
  
  downMics <- downMics %>%
  group_by(sample) %>%
  summarize(tempDown = sum(exo.ra))
  
  if(upTotal == 0 && downTotal == 0){
    df <- data.frame(matrix(ncol = 0, nrow = 0))
  }else if(upTotal == 0){
    df <- downMics %>%
      mutate(score = tempDown*(1/downTotal)) %>%
      select(sample,score)
  }else if(downTotal == 0){
    df <- upMics %>%
      mutate(score = tempUp*(1/upTotal)) %>%
      select(sample, score)
  }else{
    df <- merge(upMics, downMics, by = "sample") %>%
    mutate(score = (tempUp - tempDown)*(1/(upTotal + downTotal))) %>%
    select(sample, score)
  }
  
  return(df)
}
```

```{r}
fitness <- getMicroSigScore(paa_up, paa_down, f_taxa) 
bewell <- getMicroSigScore(paa_up, paa_down, b_taxa)
```

rename
```{r}
fitness$project <- "fitness"
bewell$project <- "bewell"

fitness <- fitness %>%
  rename("paaScore" = score)

bewell <- bewell %>%
  rename("paaScore" = score)
```



### test at genus level ###

define paa genus mics
```{r}
paa_gen_up <- c("g__Clostridium", "g__Firmicutes_unclassified", "g__Coprobacillus", "g__Streptococcus", "g__Acidaminococcus", "g__Roseburia", "g__Butyrivibrio")

paa_gen_down <- c("g__Alistipes", "g__Fusobacterium", "g__Leuconostoc", "g__Paenibacillus","g__Alcanivorax", "g__Arthobacter", "g__Bacillus")
```

```{r}
fitnessGen <- getMicroSigScore(paa_gen_up, paa_gen_down, f_taxa) 
bewellGen <- getMicroSigScore(paa_gen_up, paa_gen_down, b_taxa)
```


get paa logical


if, of microbes present in data period within paa signature, all are present within sample => TRUE else => FALSE

check microbes present in data
```{r}
flog <- f_taxa[grepl(paste(paa_up, collapse="|"), f_taxa$microbe),]
blog <- b_taxa[grepl(paste(paa_up, collapse="|"), b_taxa$microbe),]
flog_gen_up <- f_taxa[grepl(paste(paa_gen_up, collapse="|"), f_taxa$microbe),]
flog_gen_down <- f_taxa[grepl(paste(paa_gen_down, collapse="|"), f_taxa$microbe),]
blog_gen_up <-  b_taxa[grepl(paste(paa_gen_up, collapse="|"), b_taxa$microbe),]
blog_gen_down <-  b_taxa[grepl(paste(paa_gen_down, collapse="|"), b_taxa$microbe),]
```

```{r}
# fitness, species level, only s__Coprobacillus_cateniformis present
fitness <- fitness %>% 
  mutate(paaBin = ifelse(paaScore == 0, FALSE, TRUE))

blog <- spread(blog, key = microbe, value = exo.ra, fill = 0)

blog <- blog %>%
  mutate(paaBin = ifelse(apply(blog == 0, 1, any), FALSE, TRUE)) %>%
  select(sample, paaBin)
  
flog_gen <- rbind(flog_gen_down, flog_gen_up)
blog_gen <- rbind(blog_gen_down, blog_gen_up)

flog_gen <- spread(flog_gen, key = microbe, value = exo.ra, fill = 0)
blog_gen <- spread(blog_gen, key = microbe, value = exo.ra, fill = 0)

flog_gen <- flog_gen %>%
  mutate(paaBin = ifelse(apply(flog_gen == 0, 1, any), FALSE, TRUE)) %>%
  select(sample, paaBin)
  
blog_gen <- blog_gen %>%
  mutate(paaBin = ifelse(apply(blog_gen == 0, 1, any), FALSE, TRUE)) %>%
  select(sample, paaBin)
```

# Want to make this process a function, come back to
```{r}
# taxa <- f_taxa
# paa <- paa_up
# data <- fitness
# 
# data$paaBin <- TRUE
# 
# log <- taxa[grepl(paste(paa, collapse="|"), taxa$microbe),]
# mics <- as.list(log %>% select(microbe) %>% distinct())
```

combine inro one table
```{r}
flog_gen <- flog_gen %>%
  rename("paaBinGenus" = "paaBin")

fitnessGen <- merge(flog_gen, fitnessGen)

fitnessGen <- fitnessGen %>%
  rename("scoreGenus" = "score")

fitness <- merge(fitness, fitnessGen)

fitness <- fitness %>%
  select(sample, project, paaScore, paaBin, scoreGenus, paaBinGenus)

blog_gen <- blog_gen %>%
    rename("paaBinGenus" = "paaBin")

bewellGen <- merge(blog_gen, bewellGen) %>%
  rename("scoreGenus" = "score") 

bewell <- merge(bewell, bewellGen)

bewell <- merge(bewell, blog)

bewell <- bewell %>%
    select(sample, project, paaScore, paaBin, scoreGenus, paaBinGenus)

paa_table <- rbind(fitness, bewell)
```


```{r}
write.csv(paa_table, "../data/derived/paa_table.csv")
```

