---
title: "cytokines ordination"
author: "Bailey Conrad"
date: "`r Sys.Date()`"
output: html_document
---

CW changed file paths 
```{r}
cytokines <- read.csv("./data/derived/cytokines_avg-results.csv", stringsAsFactors = F)
samplabs <- readRDS("../../data/derived/clinical_manuscript.rds")$samples
#demos <- read.csv("/Users/baileyconrad/Documents/BEWELL demographics.csv")
```

CW 
```{r}
cyto.form <- cytokines %>%
  select(-X) %>% 
  pivot_wider(names_from = cytokine, values_from = avg.res)

names <- cyto.form$Sample
cyto.form <- cyto.form[,-1]
row.names(cyto.form) <- names

cyto.pca <- prcomp(na.omit(cyto.form) , center = TRUE)

PCi<-data.frame(cyto.pca$x)

PCi %>%
  ggplot(aes(x=PC1, y=PC2)) + 
           geom_point()

```



#attempting to run fxn with raw cytokines data and trying to just have numbers
```{r}
cytokines$cytokine = as.numeric(as.factor(cytokines$cytokine))
   
cytokines %>% select(-X) %>% select(-Sample)

cytokines.pca <- prcomp(cytokines)
summary(cytokines.pca)
```




```{r}
cytokines %>% select(-X) %>% pivot_wider(names_from = cytokine, values_from = avg.res)
```

```{r}
combine <- merge(samplabs,demos,by="Subject.ID")
```


```{r}
#piping manipulations until you have what you want
#all we want is sample id and berry group (labels for all 4 groups)
#we are testing post BRB to preP, postP, and preBRB
samplab.form <- combine %>%
  select(Subject.Number, placebo, Smoking.status) %>%
  mutate(preP = ifelse(placebo == "Week 0", 1, 3), #quote bc exact string #==; is equal to operator #true = 1, false =3
         postP = ifelse(placebo == "Week 0", 2, 4), 
         preBRB = ifelse(placebo == "Week 0", 3, 1),
        postBRB = ifelse(placebo == "Week 0", 4, 2)) %>%
  pivot_longer(names_to = "timepoint.label", values_to = "timepoints", cols = c("preP", "postP", "preBRB", "postBRB")) %>%
  mutate(Sample = paste0(Subject.Number, "-", timepoints)) %>%
  select(Sample, timepoint.label, Smoking.status) 
  # filter(Smoking.status=="former")


cyto.form <- cytokines %>%
 select(-X) %>% pivot_wider(names_from = cytokine, values_from = avg.res) %>%
  mutate(Sample = gsub("^0", "", Sample), Sample = gsub("\\.1$", "", Sample)) %>%
  left_join(samplab.form) %>%
  mutate(berry.group = ifelse(grepl("postBRB",timepoint.label), 1, 0)) %>%
  filter(Smoking.status=="former") %>%
  # subset(select = -timepoint.label) %>%
  # subset (select = -Smoking.status) %>%
  # pivot_longer(names_to = "timepoint.label", values_to = "cytokines", cols = c(""))
```

#Yang Yang's code below

#PCA analysis

```{r PCA of cleaned up data file}
#PCA
cytokines.pca <- prcomp(cytokines)
summary(cytokines.pca)
```
#Visualization - clustering plot

```{r Clustering plot}
#graph op1 with only labels
autoplot(seqtabc.pca, data = metamerge.match.c, colour = "GroupID", label = TRUE, label.size = 3, shape = FALSE, frame = T, frame.type = "norm") +
  theme_bw() 
#graph alt2 with only dots
colnames(metamerge.match.c)[4] <- "Experimental.treatment.groups"
clusterplot <-
autoplot(seqtabc.pca, data = metamerge.match.c, colour = "Experimental.treatment.groups", frame = T, frame.type = "norm") +
  theme_bw()+ 
  xlim(0.0, 0.25) +
  ylim(-1.5, 1.5) +
  ggtitle("Effect of BRB Nectar on Microbiome Composition", subtitle = " PCA Clustring Analysis by Experimental Groups")
save_plot(clusterplot, "../figures/Group.Clustering.Graph.png")
